BOLD Parameter Estimation Using Particle Filters

### 2010-08-11 doc/thesis/conclusion.tex
more in the conclusion

\chapter{Conclusion}
\label{sec:Conclusion}
This work has demonstrated the use of the particle filter to
learn parameters of the BOLD model. Since the inception of the
BOLD equations, many attempts have been made to learn the parameters,
given FMRI data. These attempts have typically either been extremely 
slow or made extensive approximations. While the particle filter method
is not quick, 60 seconds to analyze each voxel is well within the capabilities
of the typical research lab.

One significant finding of this work, that would not be clear without 
calculating a full posterior distribution, is the interplay 
between the parameters. The results of simulations clearly demonstrate
that identifying a single set of parameters is not possible with this 
model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
at this, the current work clearly demonstrates this fact. Therefore,
any single estimate of parameters is insufficient for analysis. As such,
approaches to the BOLD model that do not treat the parameters as distributions
will not be able to overcome the inherent indeterminability of the
parameters. Besides the Unscented Kalman Filter, this is the only approach
that accomplishes this task without recalculating the parameters over and 
over to build a distribution. The Unscented Kalman Filter, though, is limited
to Gaussian estimation.

>>>>

\chapter{Conclusion}
\label{sec:Conclusion}
This work has demonstrated the use of the particle filter to
learn parameters of the BOLD model. Since the inception of the
BOLD equations, many attempts have been made use FMRI data to 
to learn these parameters. These attempts have typically either been extremely 
slow or relied on extensive assumptions. While the particle filter method
is not quick, 60 seconds to analyze each voxel is well within the capabilities
of the typical research lab. Previous attempts have also treated
the problem as if there were a single solution.

One significant finding of this work, that would not be clear without 
calculating a full posterior distribution, is the interplay 
between the parameters. The results of simulations clearly demonstrate
that identifying a single set of parameters is not possible with this 
model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
at this, the current work clearly demonstrates this fact. Therefore,
any single estimate of parameters is insufficient for analysis. As such,
approaches to the BOLD model that do not treat the parameters as distributions
will not be able to overcome the inherent indeterminability of the
parameters. Besides the Unscented Kalman Filter, this is the only approach
that accomplishes this task without repeatedly calculating the parameters 
to build a distribution. The Unscented Kalman Filter, of course, is limited
to Gaussian estimation which limits its ability to estimate the posterior.

The primary reason this method has not been used before is the high 
dimensionality of the system. In \cite{Murray2008} the idea of learning
the parameters is floated as the correct method, yet 
learning 7 parameters with a Monte Carlo method is often intractable. The 
concern was that so many parameters creates a prohibitively large 
search space, that requires too many particle to learn properly. To overcome
this difficulty, instead of starting the algorithm with 1000 particles,
the initial number of particles was set to 16,000. This meant that
when the search space was the largest, the number of particles was
sufficiently dense to represent the prior distribution. Then resampling
was performed for the first time, the number of particles was dropped, 
since the weight of most the particles had dropped to 0. The result is
a particle filter algorithm that spends computing resources only where
they are really needed.

Many of the approaches to determining areas of neural activation aim
to be Bayesian, yet ultimately must make limiting assumptions about
the distributions that turn the approach into classical statistics. 
This approach is a true Bayesian non-parametric approach that given
enough measurements will approach the true probability distribution
of the parameters. While the conclusion that the parameters are not
uniquely identifiable is a disappointing in light years of research 
attempting to learn these parameters, it in fact further demonstrates
the need to estimate probability distribution rather than a single 
parameter estimate.
